---
title: "Baby Names"
format:
  html:
    theme: flatly
    mainfont: "Outfit, sans-serif"
    include-in-header:
      text: |
        <link rel="preconnect" href="https://fonts.googleapis.com">
        <link rel="preconnect" href="https://fonts.gstatic.com" crossorigin>
        <link href="https://fonts.googleapis.com/css2?family=Outfit:wght@300;400;600;700&display=swap" rel="stylesheet">
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
library(plotly)
```

```{r}
#| label: load-data
# Find all yearly text files and read them into one tibble
files <- list.files("data", pattern = "^yob[0-9]{4}\\.txt$", full.names = TRUE)

babynames <- read_csv(
  files,
  id = "year",
  col_names = c("name", "sex", "n")
) |>
  mutate(year = parse_number(year))

# Save aggregated table for fast interactive queries in the browser
if (!file.exists("data/names.parquet")) {
  babynames |>
    summarise(n = sum(n), .by = c(name, year)) |>
    arrow::write_parquet("data/names.parquet")
}
```

::: {.column-margin}
### Similar Names

Names most similar in spelling to **Brooklynn**:

```{r}
#| label: similar-names-table
# Calculate spelling similarity to Brooklynn using Levenshtein distance
candidates <- c(
  "Brooklyn", "Brooklynne", "Brooklyne", "Brookelynn", 
  "Brooklin", "Brookelyn", "Braylynn", "Brooke", "Oaklynn"
)

tibble(
  Name = candidates,
  Similarity = paste0(round((1 - adist("Brooklynn", candidates)[1, ] / 10) * 100), "%")
) |>
  knitr::kable()
```
:::

The rise of the spelling **Brooklynn** was largely driven by the surge of the primary spelling **Brooklyn** as a popular place-name and girl's name, especially after David and Victoria Beckham named their firstborn Brooklyn in 1999.

```{r}
#| label: plot-brooklyn-comparison
# Compare both spellings to see the shared trend and the Beckham milestone
p <- babynames |>
  filter(name %in% c("Brooklyn", "Brooklynn")) |>
  summarise(total = sum(n), .by = c(year, name)) |>
  ggplot(aes(x = year, y = total, color = name)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 1999, linetype = "dashed", color = "gray40") +
  annotate(
    "text",
    x = 1998,
    y = 6000,
    label = "1999: Brooklyn Beckham born",
    hjust = 1,
    size = 3.5,
    family = "URW Gothic"
  ) +
  labs(
    title = "Popularity of Brooklyn vs. Brooklynn Over Time",
    x = "Year",
    y = "Number of Babies",
    color = "Spelling"
  ) +
  theme_minimal(base_family = "URW Gothic", base_size = 12)

ggplotly(p)
```

## Compare Your Name

Type any name into the box below to compare its popularity to Brooklynn:

```{ojs}
//| echo: false
db = DuckDBClient.of({names: FileAttachment("data/names.parquet")})

viewof your_name = Inputs.text({
  label: "Enter a name:",
  value: "Jessica",
  placeholder: "Type a name..."
})

comparison_data = db.query(`
  SELECT year, name, n 
  FROM names 
  WHERE LOWER(name) IN ('brooklynn', LOWER('${your_name}'))
  ORDER BY year
`)

Plot.plot({
  title: `Comparing ${your_name} with Brooklynn`,
  marks: [
    Plot.lineY(comparison_data, {x: "year", y: "n", stroke: "name", strokeWidth: 2}),
    Plot.tip(comparison_data, Plot.pointerX({x: "year", y: "n", fill: "name"}))
  ],
  x: {tickFormat: "d", label: "Year"},
  y: {label: "Number of Babies", grid: true},
  color: {legend: true}
})
```


